\pkg{ShinyMICE}: an Evaluation Suite for Multiple Imputation

### 2019-11-11 ResearchReport/article.Rnw
Cleaned up article file, solved mistake in calculation CI.

With MI, many assumptions are made about the nature of the observed and missing parts of the data and their relation to the 'true' \emph{data generating model} \citep{buur18}. Without proper evaluation of the imputations and the underlying assumptions, any drawn inference may erroneously be deemed valid. Such evaluation measures are currently missing or under-developed in MI software, like the world leading \proglang{R} package \pkg{mice} \citep{mice11}. Therefore, I will answer the following question: 'Which measures are vital for evaluating the validity of multiply imputed data?'.

@@

The aim of this paper is to provide applied researchers an introduction to the interactive evaluation device \pkg{ShinyMICE}. All programming code used in this paper is available in the file XYZ.R along with the manuscript, and on Github repository XYZ.

"The intended audience of this paper consists of applied researchers who want to address problems caused by missing data by multiple imputation. The text assumes basic familiarity with R. The document contains hands-on analysis using the mice package. We do not discuss problems of incomplete data in general. We refer to the excellent books by Little and Rubin (2002) and Schafer (1997). Theory and applications of multiple imputation have been developed in Rubin (1987) and Rubin (1996). van Buuren (2012) introduces multiple imputation from an applied perspective" \cite[p.~4]{mice}.

%% -- Notation ---------------------------------------------------------------
\subsection{Notation and conventions?} \label{sec:notation}

Notation in this paper is copied from \cite{buur18}, and thus based on \cite{rubin87}. Let $Y$ denote an $n \times p$ matrix containing the data values on $p$ variables for all $n$ units in a sample. The collection of observed data values is denoted as $Y_{obs}$; the missing part of $Y$ is refered to as $Y_{mis}$. Response indicator $R$ shows whether a data value in $Y$ is missing or observed. The relation between $R$, $Y_{obs}$, and $Y_{mis}$ determines the missingness mechanism.

@@

- Robustness of inferences to varying degrees of violation can be assessed with sensitivity analyses. See \citep{nguy17}.

- Forked MICE with MNAR sensitivity analyses, see \url{https://thestatsgeek.com/2018/06/07/missing-not-at-random-sensitivity-analysis-with-fcs-multiple-imputation/}.

- FIMD on sensitivity analyses, see https://stefvanbuuren.name/fimd/sec-sensitivity.html

"The MAR assumption can never be tested from the observed data. One can however check whether the imputations created by MICE algorithm are plausible" \cite[p.~42]{mice}.

"Plot densities of both the observed and imputed values of all variables to see whether the imputations are reasonable. Differences in the densities between the observed and imputed values may suggest a problem that needs to be further checked" \cite[p.~43]{mice}.

@@

- Implement existing plot functions in \pkg{ShinyMICE}, add subsequent statistical tests to quantify the relations with covariates.

- Relations between observed and missing parts of the data: use \pkg{lattice} package functionalities to improve the \pkg{mice} functions \fct{hist} and \fct{xyplot}.

- Evaluation of MI methods in SAS, see https://onlinelibrary.wiley.com/doi/full/10.1111/1467-9574.00219

"An important step in multiple imputation is to assess whether imputations are plausible. Imputations should be values that could have been obtained had they not been missing. Imputations should be close to the data. Data values that are clearly impossible (e.g., negative counts, pregnant fathers) should not occur in the imputed data. Imputations should respect relations between variables, and reflect the appropriate amount of uncertainty about their 'true' values. Diagnostic checks on the imputed data provide a way to check the plausibility of the imputations" \cite[p.~11]{mice}.

"The mice package contains several graphic functions that can be used to gain insight into the correspondence of the observed and imputed data: \fct{bwplot}, \fct{stripplot}, \fct{densityplot} and \fct{xyplot}." \cite[par.~6.6]{buur18}

"It is often useful to inspect the distributions of original and the imputed data. One way of doing this is to use the function \fct{stripplot} \dots  The differences between the red points represents our uncertainty about the true (but unknown) values. \dots Under MCAR, univariate distributions of the observed and imputed data are expected to be identical. Under MAR, they can be different, both in location and spread, but their multivariate distribution is assumed to be identical." \cite[p.~12]{mice}.

"Bondarenko and Raghunathan (2016) [see https://www.ncbi.nlm.nih.gov/pubmed/26952693] proposed a more refined diagnostic tool that aims to compare the distributions of observed and imputed data conditional on the missingness probability. The idea is that under MAR the conditional distributions should be similar if the assumed model for creating multiple imputations has a good fit. An example is created as ..."  \cite[par.~6.6]{buur18}

% fit <- with(imp, glm(ici(imp) ~ age + bmi + hyp + chl,
%                      family = binomial))
% ps <- rep(rowMeans(sapply(fit$analyses, fitted.values)),
%           imp$m + 1)
% xyplot(imp, bmi ~ ps | as.factor(.imp),
%        xlab = "Probability that record is incomplete",
%        ylab = "BMI", pch = c(1, 19), col = mdc(1:2))

@@

- Convergence issues: "Imputers who do choose to use FCS should use flexible univariate models wherever possible and take care to assess apparent convergence of the algorithm, for example by computing traces of pooled estimates or other statistics and using standard MCMC diagnostics (Gelman et al., 2013, Chapter 11). It may also be helpful to examine the results of many independent runs of the algorithm with different initializations and to use random scans over the p variables to try to identify any convergence issues and mitigate possible order dependence" \cite[p.~19]{murr18}.

- Replicate simulation study and build a decision rule to solve the problem: "The monitoring statistic was computed for mean monthly earnings at each iteration in chains of length k = 200. Since calculation of the statistic requires M parallel sequences, m = 5 such chains were constructed. This value of m was informed by the preferred choice given in the literature on multiple imputation. The monitoring statistic computed at each iteration is presented in Figures 15 to 17 for each of the three missingness mechanisms. The red vertical line denotes ten iterations" \cite[p.~49]{lace07}.

- Potentially add something about updated (2019) version of $\widehat{R}$ and the new threshold of 1.01, see \citep{veht19}.

- Auto-correlation: "Applications of MICE with lowly correlated data therefore inject a lot of noise into the system. Hence, the auto-correlation over t will be low, and convergence will be rapid, and in fact immediate if all variables are independent. Thus, the incorporation of noise into the imputed data has pleasant side-effect of speeding up convergence" \citep{buur18}, par. 4.5. Also, schafer (1997, p. 129) wrote on worst linear statistic. We could calculate the autocorrelation of that statistic to know that the algorithm converged elsewhere too. See autocorr function plot in sas of worst linear function. Note: we're talking about missing data only, not the combined data (that autocorrelation is very high, as is the autocorrelation of deductively imputed values, like in the texp example). Worst linear function in SAS see \url{https://support.sas.com/documentation/cdl/en/statug/63033/HTML/default/viewer.htm#statug_mi_sect027.htm}.

- Stability of the solution: possibly use the slope of means over iterations too to see whether there is trending. Or apply PCA on the imputed data and if that (the eigenvalues?) stays the same we know that the means and variances are stable as well. Or look at MC error: MC error = SD/sqrt(number of iterations), where SD represents the variation across iterations. The MC error thus represents how much the means differ w.r.t. the iterations. MC error decreases as number of iterations increases. It should not be larger than 5\% of the sample standard deviation.

"In general, you cannot know for sure if your chain has converged. But sometimes you can know if your chain has not converged, so we at least check for this latter possibility" \cite[p.~101]{hoff09}

"In order to converge to a stationary distribution, a Markov chain needs to satisfy three important properties (Roberts 1996; Tierney 1996): irreducible, the chain must be able to reach all interesting parts of the state space; aperiodic, the chain should not oscillate between different states; recurrence, all interesting parts can be reached infinitely often, at least from almost all starting points. Do these properties hold for the MICE algorithm? Irreducibility is generally not a problem since the user has large control over the interesting parts of the state space. This flexibility is actually the main rationale for FCS instead of a joint model. Periodicity is a potential problem, and can arise in the situation where imputation models are clearly inconsistent. A rather artificial example of an oscillatory behavior occurs when Y 1 is imputed by Y 2 $\beta$ + $\epsilon$  1 and Y 2 is imputed by - Y 1 $\beta$ + $\epsilon$  2 for some fixed, nonzero $\beta$ . The sampler will oscillate between two qualitatively different states, so the correlation between Y 1 and Y 2 after imputing Y 1 will differ from that after imputing Y 2 . In general, we would like the statistical inferences to be independent of the stopping point. A way to diagnose the ping-pong problem, or order effect, is to stop the chain at different points. The stopping point should not affect the statistical inferences. The addition of noise to create imputations is a safeguard against periodicity, and allows the sampler to "break out" more easily. Non-recurrence may also be a potential difficulty, manifesting itself as explosive or non-stationary behavior. For example, if imputations are made by deterministic functions, the Markov chain may lock up. Such cases can sometimes be diagnosed from the trace lines of the sampler. See Section 6.5.2 for an example. As long as the parameters of imputation models are estimated from the data, non-recurrence is mild or absent. The required properties of the MCMC method can be translated into conditions on the eigenvalues of the matrix of transition probabilities (MacKay 2003, 372-73). The development of practical tools that put these conditions to work for multiple imputation is still an ongoing research problem." \cite[par.~4.5]{buur18}

@@

R hat not working on MI data? "if over-dispersion does not hold, $var_plus$ can be too low, which can lead to falsely diagnosing convergence.)" \cite[p~437]{broo98}. Aim to replicate Lacerda et al. that R hat will not be smaller than 1.1 before iteration nr. 50.

@@

@@

Set-up: simulate data, impute missing data points with varying number of iterations. Then evaluate the regular diagnostics of MI simulations, see Vink, n.d., and the convergence. Diagnostics include absolute bias of the estimated regression coefficient, confidence interval width, and empirical coverage rate (coverage probability?) across simulations. Convergence is evaluated with $\hat{R}$, or potential scale reduction factor. Proposed by Gelman and Rubin (1992), but since developed into ...

@@

"The mixture-of-sequences variance, V, should stabilize as a function of n. (Before convergence, we expect $\sigma^2$ to decrease with n, only increasing if the sequences explore a new area of parameter space, which would imply that the original sequences were not overdispersed for the particular scalar summary being monitored.)" 
% \cite[p~438]{broo98}

We would like to have convergence measures for multivariable statistics (scalars?) of interest. This is, however, dependent of the complete data model. The eigenvector decomposition methof proposed by May 9? should be implemented. I could not find any resources to apply this method and it is outside the scope of this thesis to investigate how this apprach could be implemented.

@@

\section{Models and software} \label{sec:models}

\emph{This is a placeholder.}

@@

\section{Illustrations} \label{sec:illustrations}

\emph{This is a placeholder.}

@@

\section{Summary and discussion} \label{sec:summary}

\emph{This is a placeholder.}

@@

The results in this paper were obtained using \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the \pkg{mice}~\Sexpr{packageVersion("mice")} package. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.

@@

This technical paper is written by the sole author (Hanne Oberman, BSc.), with guidance from Master thesis supervisors prof. dr. Stef van Buuren, and dr. Gerko Vink.

@@

\section{More technical details} \label{app:technical}

@@

\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

>>>>

With MI, many assumptions are made about the nature of the observed and missing parts of the data and their relation to the 'true' \emph{data generating model} \citep{buur18}. Without proper evaluation of the imputations and the underlying assumptions, any drawn inference may erroneously be deemed valid. Such evaluation measures are currently missing or under-developed in MI software, like the world leading \proglang{R} package \pkg{mice} \citep{mice}. Therefore, I will answer the following question: 'Which measures are vital for evaluating the validity of multiply imputed data?'.

@@

The aim of this paper is to provide applied researchers an introduction to the interactive evaluation device \pkg{ShinyMICE}. The intended audience consists of empirical researchers and statisticians who (want to) use multiple imputation to solve missing data problems. Basic familiarity with multiple imputation methodology is assumed. For an accessible and comprehensive introduction to MI from an applied perspective, see \cite{buur18}. For the theoretical foundation of MI, see \cite{rubin87}. All programming code used in this paper is available in the file 'XYZ.R' along with the manuscript, and on Github repository 'XYZ'.

% "The intended audience of this paper consists of applied researchers who want to address problems caused by missing data by multiple imputation. The text assumes basic familiarity with R. The document contains hands-on analysis using the mice package. We do not discuss problems of incomplete data in general. We refer to the excellent books by Little and Rubin (2002) and Schafer (1997). Theory and applications of multiple imputation have been developed in Rubin (1987) and Rubin (1996). van Buuren (2012) introduces multiple imputation from an applied perspective" \cite[p.~4]{mice}.

%% -- Notation ---------------------------------------------------------------
\subsection{Notation} \label{sec:notation}

The \proglang{R} \pkg{Shiny} application introduced in this paper is developed with the aim to integrate it into the \pkg{mice} environment. \pkg{ShinyMICE} therefore follows notation and conventions of \cite{mice}. For an overview of the deviations from the 'original' notation by \cite{rubin87}, see \cite{buur18}.

Let $Y$ denote an $n \times p$ matrix containing the data values on $p$ variables for all $n$ units in a sample. The collection of observed data values is denoted as $Y_{obs}$; the missing part of $Y$ is refered to as $Y_{mis}$. Response indicator $R$ shows whether a data value in $Y$ is missing or observed. The relation between $R$, $Y_{obs}$, and $Y_{mis}$ determines the missingness mechanism.

@@

% In general: "The MAR assumption can never be tested from the observed data. One can however check whether the imputations created by MICE algorithm are plausible" \cite[p.~42]{mice}.

\subsubsection{What is already implemented?}

- Current guideine: "Plot densities of both the observed and imputed values of all variables to see whether the imputations are reasonable. Differences in the densities between the observed and imputed values may suggest a problem that needs to be further checked" \cite[p.~43]{mice}. 
% FIMD on sensitivity analyses, see https://stefvanbuuren.name/fimd/sec-sensitivity.html

\subsubsection{What is not yet implemented, but exists?}

- Robustness of inferences to varying degrees of violation can be assessed with sensitivity analyses. See \citep{nguy17} for guidelines.

- Forked MICE with MNAR sensitivity analyses
% see \url{https://thestatsgeek.com/2018/06/07/missing-not-at-random-sensitivity-analysis-with-fcs-multiple-imputation/}.

\subsubsection{What is not implemented, and NA?}

- Nothing within the scope of this thesis.

@@

% In general: "An important step in multiple imputation is to assess whether imputations are plausible. Imputations should be values that could have been obtained had they not been missing. Imputations should be close to the data. Data values that are clearly impossible (e.g., negative counts, pregnant fathers) should not occur in the imputed data. Imputations should respect relations between variables, and reflect the appropriate amount of uncertainty about their 'true' values. Diagnostic checks on the imputed data provide a way to check the plausibility of the imputations" \cite[p.~11]{mice}.

\subsubsection{What is already implemented?}

- Existing plot functions to evaluate relations between observed and missing parts of the data (\pkg{mice} functions \fct{hist}, \fct{xyplot}, and \fct{stripplot}). 
% "The mice package contains several graphic functions that can be used to gain insight into the correspondence of the observed and imputed data: \fct{bwplot}, \fct{stripplot}, \fct{densityplot} and \fct{xyplot}." \cite[par.~6.6]{buur18}

% "It is often useful to inspect the distributions of original and the imputed data. One way of doing this is to use the function \fct{stripplot} \dots  The differences between the red points represents our uncertainty about the true (but unknown) values. \dots Under MCAR, univariate distributions of the observed and imputed data are expected to be identical. Under MAR, they can be different, both in location and spread, but their multivariate distribution is assumed to be identical." \cite[p.~12]{mice}.

\subsubsection{What is not yet implemented, but exists?}

- Improve plots using \pkg{lattice} package functionalities.

- Subsequent statistical tests to quantify the relations with covariates.
% "Bondarenko and Raghunathan (2016) [see https://www.ncbi.nlm.nih.gov/pubmed/26952693] proposed a more refined diagnostic tool that aims to compare the distributions of observed and imputed data conditional on the missingness probability. The idea is that under MAR the conditional distributions should be similar if the assumed model for creating multiple imputations has a good fit. An example is created as ..."  \cite[par.~6.6]{buur18}
% 
% fit <- with(imp, glm(ici(imp) ~ age + bmi + hyp + chl,
%                      family = binomial))
% ps <- rep(rowMeans(sapply(fit$analyses, fitted.values)),
%           imp$m + 1)
% xyplot(imp, bmi ~ ps | as.factor(.imp),
%        xlab = "Probability that record is incomplete",
%        ylab = "BMI", pch = c(1, 19), col = mdc(1:2))

- Use evaluation of MI methods in SAS as example
% see https://onlinelibrary.wiley.com/doi/full/10.1111/1467-9574.00219

\subsubsection{What is not implemented, and NA?}

- Nothing within the scope of this thesis.

@@

% In general: "In general, you cannot know for sure if your chain has converged. But sometimes you can know if your chain has not converged, so we at least check for this latter possibility" \cite[p.~101]{hoff09}

% "In order to converge to a stationary distribution, a Markov chain needs to satisfy three important properties (Roberts 1996; Tierney 1996): irreducible, the chain must be able to reach all interesting parts of the state space; aperiodic, the chain should not oscillate between different states; recurrence, all interesting parts can be reached infinitely often, at least from almost all starting points. Do these properties hold for the MICE algorithm? Irreducibility is generally not a problem since the user has large control over the interesting parts of the state space. This flexibility is actually the main rationale for FCS instead of a joint model. Periodicity is a potential problem, and can arise in the situation where imputation models are clearly inconsistent. A rather artificial example of an oscillatory behavior occurs when Y 1 is imputed by Y 2 $\beta$ + $\epsilon$  1 and Y 2 is imputed by - Y 1 $\beta$ + $\epsilon$  2 for some fixed, nonzero $\beta$ . The sampler will oscillate between two qualitatively different states, so the correlation between Y 1 and Y 2 after imputing Y 1 will differ from that after imputing Y 2 . In general, we would like the statistical inferences to be independent of the stopping point. A way to diagnose the ping-pong problem, or order effect, is to stop the chain at different points. The stopping point should not affect the statistical inferences. The addition of noise to create imputations is a safeguard against periodicity, and allows the sampler to "break out" more easily. Non-recurrence may also be a potential difficulty, manifesting itself as explosive or non-stationary behavior. For example, if imputations are made by deterministic functions, the Markov chain may lock up. Such cases can sometimes be diagnosed from the trace lines of the sampler. See Section 6.5.2 for an example. As long as the parameters of imputation models are estimated from the data, non-recurrence is mild or absent. The required properties of the MCMC method can be translated into conditions on the eigenvalues of the matrix of transition probabilities (MacKay 2003, 372-73). The development of practical tools that put these conditions to work for multiple imputation is still an ongoing research problem." \cite[par.~4.5]{buur18}

\subsubsection{What is already implemented?}

- Traceplots

\subsubsection{What is not yet implemented, but exists?}

- $\widehat{R}$, but too stringent (new) threshold, and assumption of overdispersed initial values of imputation chains not met.
% Potentially add something about updated (2019) version of $\widehat{R}$ and the new threshold of 1.01, see \citep{veht19}.

- Autocorrelation. Schafer (1997, p. 129) wrote on worst linear statistic. We could calculate the autocorrelation of that statistic to know that the algorithm converged elsewhere too. See autocorr function plot in sas of worst linear function.
% Worst linear function in SAS see \url{https://support.sas.com/documentation/cdl/en/statug/63033/HTML/default/viewer.htm#statug_mi_sect027.htm}.
% Note: we're talking about missing data only, not the combined data (that autocorrelation is very high, as is the autocorrelation of deductively imputed values, like in the texp example). 
% "Applications of MICE with lowly correlated data therefore inject a lot of noise into the system. Hence, the auto-correlation over t will be low, and convergence will be rapid, and in fact immediate if all variables are independent. Thus, the incorporation of noise into the imputed data has pleasant side-effect of speeding up convergence" \citep{buur18}, par. 4.5.

- Sensitivity analysis: Run algorithm several times and compare results. 
% "Imputers who do choose to use FCS should use flexible univariate models wherever possible and take care to assess apparent convergence of the algorithm, for example by computing traces of pooled estimates or other statistics and using standard MCMC diagnostics (Gelman et al., 2013, Chapter 11). It may also be helpful to examine the results of many independent runs of the algorithm with different initializations and to use random scans over the p variables to try to identify any convergence issues and mitigate possible order dependence" \cite[p.~19]{murr18}.

\subsubsection{What is not implemented, and NA?}

- $\hat{R}$ threshold: Replicate simulation study and build a decision rule to solve the problem with $\hat{R}$.
% "The monitoring statistic was computed for mean monthly earnings at each iteration in chains of length k = 200. Since calculation of the statistic requires M parallel sequences, m = 5 such chains were constructed. This value of m was informed by the preferred choice given in the literature on multiple imputation. The monitoring statistic computed at each iteration is presented in Figures 15 to 17 for each of the three missingness mechanisms. The red vertical line denotes ten iterations" \cite[p.~49]{lace07}.

- Stability of the solution: Possibly use the slope of means over iterations too to see whether there is trending. Or apply PCA on the imputed data and if that (the eigenvalues?) stays the same we know that the means and variances are stable as well, see McKay (?).

- MC error: MC error = SD/sqrt(number of iterations), where SD represents the variation across iterations. The MC error thus represents how much the means differ w.r.t. the iterations. MC error decreases as number of iterations increases. It should not be larger than 5\% of the sample standard deviation.

@@

\emph{It is written as a (mini) thesis, with an introduction, methods section, some results (i.e., preliminary analyses, or pilot simulations), and a discussion of results. The length of the research report should be maximally 2500 words of text (without references list and or tables and figures). Please do not include appendices, and no more than 6 tables or figures. Table and Figure captions do not count towards the word limit. An abstract may be included, but is not necessary.}

R hat not working on MI data? "if over-dispersion does not hold, $\sigma_+^2$ can be too low, which can lead to falsely diagnosing convergence." \cite[p~437]{broo98}. Aim to replicate Lacerda et al. that R hat will not be smaller than 1.1 before iteration nr. 50.

@@

Evaluate: bias, variance, and cov rate to find problems that you might miss otherwise. There is no tradeoff between them. Also look at distributional char., and plausibility of imp. values, see \cite{vinknd}.

% Distributional characteristics: In practice, the distribution of the incomplete data may differ greatly from the observed data. Under anything but the MCAR assumption, this can be expected. When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed (see Abayomi et al., 2008, for an detailed overview of diagnostic evaluation for multivariate imputations). When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.

% Plausibility of the imputed values: Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference. However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property. When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible value

@@

Set-up: simulate data, impute missing data points with varying number of iterations. Then evaluate the regular diagnostics of MI simulations, see Vink, n.d., and the convergence. Diagnostics include absolute bias of the estimated regression coefficient, confidence interval width, and empirical coverage rate (coverage probability?) across simulations. Convergence is evaluated with $\hat{R}$, or potential scale reduction factor. Proposed by Gelman and Rubin (1992), but since developed into ...
% Missingness mechanism: "With MCAR missingness mechanisms, the probability to be missing is the same for all cases. This is a necessary simulation condition for evaluating the performance of imputation procedures. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not sound." (Vink, n.d., p. 4)

@@

"The mixture-of-sequences variance, V, should stabilize as a function of n. (Before convergence, we expect $\sigma^2$ to decrease with n, only increasing if the sequences explore a new area of parameter space, which would imply that the original sequences were not overdispersed for the particular scalar summary being monitored.)" \cite[p~438]{broo98}.

We would like to have convergence measures for multivariable statistics (scalars?) of interest. This is, however, dependent of the complete data model. The eigenvector decomposition methof proposed by McKay (?) should be implemented. I could not find any resources to apply this method and it is outside the scope of this thesis to investigate how this apprach could be implemented.

@@

% \section{Models and software} \label{sec:models}
% 
% \emph{This is a placeholder.}

@@

% \section{Illustrations} \label{sec:illustrations}
% 
% \emph{This is a placeholder.}

@@

% \section{Summary and discussion} \label{sec:summary}
% 
% \emph{This is a placeholder.}

@@

The results in this paper were obtained using \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the \pkg{mice}~\Sexpr{paste(packageVersion("mice"), collapse = ".")} package. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.

@@

This paper is written by the sole author (Hanne Oberman, BSc.), with guidance from Master thesis supervisors prof. dr. Stef van Buuren, and dr. Gerko Vink.

@@

% \section{More technical details} \label{app:technical}

@@

% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

### 2019-11-11 ResearchReport/article.tex
Cleaned up article file, solved mistake in calculation CI.

@@

\begin{Code}
# pseudocode of similation 
simulate data with missingness
for (number of iterations from 1 to 100)
  impute the missingness
  compute convergence and simulation diagnostics
save diagnostics
\end{Code}

Simulation code is available in appendix A and on Github.

>>>>

bias, variance, cov rate: evaluate problems that you might miss otherwise. there is no tradeoff between them. then your inference is valid. generate data, create miss, usually there is a new method that you want to investigate and then compare to cca.

% Missingness mechanism: "With MCAR missingness mechanisms, the probability to be missing is the same for all cases. This is a necessary simulation condition for evaluating the performance of imputation procedures. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not sound." (Vink, n.d., p. 4)

@@

\begin{Code}
# pseudocode of similation 
simulate data with missingness
for (number of iterations from 1 to 100)
  impute the missingness
  compute convergence diagnostic
  perform analysis
  pool results
  compute simulation diagnostics
save diagnostics
\end{Code}

Simulation code is available in online appendix XYZ on Github.
